---
output:
  html_document: default
  pdf_document: default
---

The code for calculating lexical category biases is based on:
https://github.com/langcog/wordbank-book/blob/master/011-categories-syntactic.Rmd

You will also need the dependencies in:
https://github.com/langcog/wordbank-book/blob/master/_common.R


```{r}
source("_common.R")
```


# Prep

```{r catsyn-items}
items <- items %>%
  filter(item_kind == "word") %>%
  mutate(num_item_id = as.numeric(substr(item_id, 6, nchar(item_id))))

items_summary <- items %>% group_by(language, form, lexical_category, lexical_class) %>% 
  summarize(N.item=n())
```

```{r pressure1, echo=FALSE}
get_vocab_comp_data <- function(input_language, input_form, input_form_type) {
  print(paste(input_language, input_form, input_form_type))
  
  language_items <- items %>%
      filter(language == input_language, form_type == input_form_type,
             lexical_category %in% c("nouns", "predicates", "function_words"))
  
  language_instrument <- get_instrument_data(input_language, input_form) %>%
    filter(value == "produces" | value == "understands")
  
  num_words <- nrow(language_items)
  each_num_words <- language_items %>%
    group_by(lexical_category) %>%
    summarise(total = n()) 

  language_comp <- left_join(language_instrument, language_items, by = "item_id", relationship = "many-to-many") %>%
    rename(measure = value) %>%
    filter(!is.na(lexical_category)) %>% # null lexical categories means they are not nouns, predicates, or function words
    group_by(data_id, measure, lexical_category) %>%
    summarise(num_true = n(), .groups = "drop") %>%
    complete(data_id, measure, lexical_category, fill = list(num_true = 0)) %>%
    left_join(each_num_words, by = "lexical_category")
  
  language_summary <- language_comp %>%
    group_by(data_id, measure) %>%
    mutate(vocab_num = sum(num_true), vocab = vocab_num/num_words, prop = num_true/total, prop_vocab=prop/sum(prop)) %>%
    ungroup()
}
```

```{r pressure2, echo=FALSE}
get_class_data <- function(input_language, input_form, input_form_type) {
  print(paste(input_language, input_form, input_form_type))
  
  language_items <- items %>%
      filter(language == input_language, form_type == input_form_type,
             lexical_category %in% c("nouns", "predicates", "function_words"))
  
  language_instrument <- get_instrument_data(input_language, input_form) %>%
    filter(!is.na(value) & value != "")

  num_words <- nrow(language_items)
  each_num_words <- language_items %>%
    group_by(lexical_class) %>%
    summarise(total = n()) 

  language_comp <- left_join(language_instrument, language_items, by = "item_id", relationship = "many-to-many") %>%
    rename(measure = value) %>%
    filter(!is.na(lexical_class)) %>% # null lexical categories means they are not nouns, predicates, or function words
    group_by(data_id, measure, lexical_class) %>%
    summarise(num_true = n(), .groups = "drop") %>%
    complete(data_id, measure, lexical_class, fill = list(num_true = 0)) %>%
    left_join(each_num_words, by = "lexical_class")
  
  language_summary <- language_comp %>%
    group_by(data_id, measure) %>%
    mutate(vocab_num = sum(num_true), vocab = vocab_num/num_words, prop = num_true/total, prop_vocab=prop/sum(prop)) %>%
    ungroup()
}
```

```{r catsyn-vocab_comp, eval=FALSE}
instruments_inc <- instruments %>%
  filter(form_type %in% c(WSs, WGs)) %>%
  select(language, form, form_type) %>%
  distinct()

# This takes a while... (no need to run if you already have the vocab_comp_data.feather)
vocab_comp_data <- instruments_inc %>%
  mutate(inst_vocab_data = pmap(list(language, form, form_type), get_vocab_comp_data)) %>%
  unnest(inst_vocab_data)

class_data <- instruments_inc %>%
  mutate(inst_vocab_data = pmap(list(language, form, form_type), get_class_data)) %>%
  unnest(inst_vocab_data)

admin_data <- pmap_dfr(instruments_inc,
  \(language, form, form_type) {
    get_administration_data(language = language, form = form, include_demographic_info = TRUE ) %>%
      select(data_id, age, language, form, child_id, sex) %>%
      mutate(form_type = form_type, .after = form)})


write_feather(vocab_comp_data, "data/categories-syntactic/vocab_comp_data.feather")
write_feather(class_data, "data/categories-syntactic/class_data.feather")
write_feather(admin_data, "data/categories-syntactic/admin_data.feather")
```

```{r catsyn-sample-sizes}
vocab_comp_data <- read_feather("data/categories-syntactic/vocab_comp_data.feather") %>%
  mutate(lexical_category = lexical_category %>%
           fct_relevel(c("nouns", "predicates", "function_words")) %>%
           fct_recode("Nouns" = "nouns", "Predicates" = "predicates",
                      "Function words" = "function_words"),
         measure = fct_relevel(measure, "understands")) %>%
  group_by(language, form_type) %>%
  mutate(num_forms = n_distinct(form),
         langform = if_else(num_forms > 1,
                            paste(language, form, sep = .inst_sep), language))

class_data <- read_feather("data/categories-syntactic/class_data.feather")
admin_data <- read_feather("data/categories-syntactic/admin_data.feather")
```

```{r catsyn-plot-area-demo, fig.cap="For American English WS data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", fig.height=3}
demo_lang <- "English (American)"
# demo_lang <- "Arabic (Saudi)"
demo_form <- "WS"
demo_data <- vocab_comp_data %>% filter(language == demo_lang, form == demo_form)

pts <- seq(0, 1, 0.01)

# note: clm is from https://github.com/langcog/langcog-package/blob/master/R/clm.R
# demo_predictions <- demo_data %>%
#   group_by(lexical_category) %>%
#   nest() %>%
#   mutate(model = map(data, ~clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = .)),
#          predictions = map(model, ~broom::augment(., newdata = tibble(vocab = pts)))) %>%
#   select(lexical_category, predictions) %>%
#   unnest() %>%
#   rename(prop = .fitted)

demo_predictions <- demo_data %>%
  group_by(lexical_category) %>%
  nest() %>%
  mutate(model = map(data, ~clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = .)),
         predictions = map(model, ~tibble(vocab = pts, prop = predict(., newdata = tibble(vocab = pts))))) %>%
  select(lexical_category, predictions) %>%
  unnest(predictions)

demo_diagonal <- cross_df(list(vocab = rev(pts),
                               lexical_category = unique(demo_data$lexical_category))) %>%
  mutate(prop = vocab)

demo_area_poly <- bind_rows(demo_predictions, demo_diagonal)

vocab_scale <- scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                  breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                  name = "Vocabulary size")
prop_scale <- scale_y_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                 breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                 name = "Proportion of category")

ggplot(demo_predictions, aes(x = vocab, y = prop, colour = lexical_category)) +
  facet_grid(. ~ lexical_category) +
  coord_fixed() +
  geom_point(data = demo_data, alpha = .03, size = 0.6) +
  # geom_point(data = demo_data, alpha = .1, size = 0.8) +
  geom_line(size = 1.5) +
  geom_polygon(data = demo_area_poly, aes(fill = lexical_category), alpha = 0.2) +
  vocab_scale +
  prop_scale +
  .scale_colour_discrete(guide = FALSE) +
  .scale_fill_discrete(guide = FALSE)+
  theme_bw()
```

```{r catsyn-resample}
# note: clm is from https://github.com/langcog/langcog-package/blob/master/R/clm.R
poly_area <- function(group_data) {
  model <- tryCatch(
    clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = group_data),
    error = function(e) return(NA)
  )
  if (all(is.na(model))) return(NA)
  return((model$solution %*% c(1/4, 1/3, 1/2) - 0.5)[1])
}

sample_areas <- function(vocab_data, group, nboot = 1000) {
  
  group <- rlang::enquo(group)
  
  sample_area <- function(i) {
    print(i)
    vocab_data %>%
      group_by(language, form, form_type, measure) %>%
      sample_frac(replace = TRUE) %>%
      group_by(language, form, form_type, measure, !!group) %>%
      nest() %>%
      mutate(area = map_dbl(data, poly_area),
             sample = i) %>%
      select(-data)
  }
  map_df(1:nboot, sample_area)
}
```

```{r catsyn-areas_compute, eval=FALSE, include=FALSE}
# vocab_comp_data <- na.omit(vocab_comp_data)
areas <- sample_areas(vocab_comp_data, group = lexical_category)
write_feather(areas, "data/categories-syntactic/vocab_comp_areas.feather")

class_areas <- sample_areas(class_data, group = lexical_class)
write_feather(class_areas, "data/categories-syntactic/class_areas.feather")
```

```{r catsyn-areas_read}
areas <- read_feather("data/categories-syntactic/vocab_comp_areas.feather")
class_areas <- read_feather("data/categories-syntactic/class_areas.feather")
```

```{r catsyn-area_summary}
area_summary <- areas %>%
  filter(!is.na(area)) %>%
  group_by(language, form, form_type, measure, lexical_category) %>%
  summarise(mean = mean(area),
            ci_lower = ci_lower(area),
            ci_upper = ci_upper(area)) %>%
  group_by(language, form_type) %>%
  mutate(num_forms = n_distinct(form),
         langform = if_else(num_forms > 1,
                            paste(language, form, sep = .inst_sep), language)) %>%
  ungroup()


lexical_list = c("nouns", "predicates", "function_words")

pred_classes <- items %>%
  filter(lexical_category %in% lexical_list) %>%
  pull(lexical_class) %>%
  unique()

pred_area_summary <- class_areas %>%
  filter(!is.na(area)) %>%
  filter(lexical_class %in% pred_classes) %>%
  group_by(language, form, form_type, measure, lexical_class) %>%
  summarise(mean = mean(area),
            ci_lower = ci_lower(area),
            ci_upper = ci_upper(area)) %>%
  group_by(language, form_type) %>%
  mutate(num_forms = n_distinct(form),
         langform = if_else(num_forms > 1,
                            paste(language, form, sep = .inst_sep), language)) %>%
  ungroup()
```

```{r catsyn-base-plot}
lexcat_colours <- set_names(.pal()(3), levels(vocab_comp_data$lexical_category))

base_plot <- function(input_forms, input_measure) {
  vocab_comp_data %>%
    filter(form %in% input_forms, measure == input_measure) %>%
    ggplot(aes(x = vocab, y = prop, colour = lexical_category)) +
      facet_wrap(~language) +
      coord_fixed() +
      vocab_scale +
      prop_scale +
      scale_colour_manual(values = lexcat_colours, name = "") +
      theme(legend.position = "top",
            legend.key = element_blank(),
            legend.background = element_rect(fill = "transparent"), 
            strip.text.x = element_text(size = 7))
}
```

# Comprehension (WG) - if needed

```{r catsyn-plot-points-wg-comp, fig.asp=1.3, fig.cap="For each language's comprehension data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}
base_plot(WGs, "understands") +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)
```

```{r catsyn-plot-areas-wg, fig.cap="Relative representation in vocabulary compared to chance for nouns and predicates for comprehension data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
plot_areas_wg <- area_summary %>%
  filter(form %in% WGs, measure == "understands") %>%
  arrange(lexical_category, mean) %>%
  mutate(order = factor(row_number()))

plot_areas <- function(area_data, measure) {
  ggplot(area_data, aes(y = order, x = mean, colour = lexical_category)) +
    ggstance::geom_pointrangeh(aes(xmin = ci_lower, xmax = ci_upper)) +
    facet_wrap(~lexical_category, scales = "free_y", drop = TRUE) +
    geom_vline(xintercept = 0, linetype = .refline, colour = .grey) +
    scale_colour_manual(values = lexcat_colours, guide = FALSE) +
    scale_y_discrete(breaks = area_data$order,
                     labels = area_data$langform) +
    labs(x = glue("Relative representation ({measure})"), y = "") +
    theme(panel.grid.major.y = .coef_line,
          axis.text.y = element_text(size = rel(0.8)))
}

plot_areas_wg %>%
  filter(lexical_category != "Function words") %>%
  plot_areas("comprehension")

plot_areas_wg_cats <- plot_areas_wg %>%
  filter(lexical_category != "Function words") %>%
  select(language, form, lexical_category, mean) %>%
  pivot_wider(names_from = lexical_category, values_from = mean)

npcor_wg <- cor.test(plot_areas_wg_cats$Nouns, plot_areas_wg_cats$Predicates)
```

```{r catsyn-function-words-comp, fig.cap="Relative representation in vocabulary compared to chance for function words for comprehension data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
 plot_areas_wg %>%
  filter(lexical_category == "Function words") %>%
  plot_areas("comprehension")
```

```{r catsyn-plot-points-ws, eval=FALSE, include=FALSE}
lexcat_colours_class <- set_names(.pal()(5), levels(class_data$lexical_class))

df_noun <- class_data %>%
  filter(lexical_class == "nouns") %>%
  select(data_id, prop, language, form, form_type, measure) %>%
  rename(vocab = prop)

df_other <- class_data %>%
  filter(lexical_class != "nouns") %>%
  select(data_id, prop, lexical_class, measure) # %>%
#  rename(prop_other = prop)

# Merge by data_id
df_merged <- inner_join(df_noun, df_other, by = c("data_id", "measure"))

# Scatter plots
noun_plot <- function(input_forms, input_measure, df_data) {
  df_data %>%
    filter(form %in% input_forms, measure == input_measure) %>%
    ggplot(aes(x = vocab, y = prop, colour = lexical_class)) +
      facet_wrap(~language) +
      coord_fixed() +
      vocab_scale +
      prop_scale +
      scale_colour_manual(values = lexcat_colours_class, name = "") +
      theme(legend.position = "top",
            legend.key = element_blank(),
            legend.background = element_rect(fill = "transparent"), 
            strip.text.x = element_text(size = 7))
}

# This takes a while...
noun_areas <- sample_areas(df_merged, group = lexical_class)

noun_area_summary <- noun_areas %>%
  filter(!is.na(area)) %>%
  # filter(lexical_class %in% pred_classes) %>%
  group_by(language, form, form_type, measure, lexical_class) %>%
  summarise(mean = mean(area),
            ci_lower = ci_lower(area),
            ci_upper = ci_upper(area)) %>%
  group_by(language, form_type) %>%
  mutate(num_forms = n_distinct(form),
         langform = if_else(num_forms > 1,
                            paste(language, form, sep = .inst_sep), language)) %>%
  ungroup()
```


## Ref = Total vocab size

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}
noun_plot(WGs, "understands", class_data) +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)
```


## Ref = Noun vocab size

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}
noun_plot(WGs, "understands", df_merged) +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)

```


## Bias strength (Ref = Total vocab size)

```{r catsyn-plot-areas-preds, fig.cap="Relative representation in vocabulary compared to chance for verbs and adjectives for production data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
plot_pred_areas <- pred_area_summary %>%
  filter(form_type == "WG", measure == "understands",
         lexical_class %in% c("verbs", "adjectives", "function_words", "nouns")) %>%
  ungroup() %>%
  arrange(lexical_class, mean) %>%
  mutate(order = factor(row_number()),
         lexical_class = lexical_class %>%
           fct_recode("Verbs" = "verbs", "Adjectives" = "adjectives", "Function Words" = "function_words", "Nouns" = "nouns") %>%
           fct_relevel("Nouns", "Verbs", "Adjectives", "Function Words"))

ggplot(plot_pred_areas, aes(y = order, x = mean)) +
  ggstance::geom_pointrangeh(aes(xmin = ci_lower, xmax = ci_upper),
                             colour = lexcat_colours["Predicates"]) +
  facet_wrap(~lexical_class, scales = "free_y", drop = TRUE, nrow=1) +
  geom_vline(xintercept = 0, linetype = .refline, colour = .grey) +
  scale_y_discrete(breaks = plot_pred_areas$order,
                   labels = plot_pred_areas$langform) +
  labs(x = glue("Relative representation (production)"), y = "") +
  theme(panel.grid.major.y = .coef_line,
        axis.text.y = element_text(size = rel(0.8)))

write.csv(plot_pred_areas, file="bias_WG_Ref_TotalVocab_2025Jun18.csv", row.names=FALSE)
```


## Bias strength (Ref = Noun vocab size)

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}

plot_pred_areas <- noun_area_summary %>%
  filter(form_type == "WG", measure == "understands",
         lexical_class %in% c("verbs", "adjectives", "function_words")) %>%
  ungroup() %>%
  arrange(lexical_class, mean) %>%
  mutate(order = factor(row_number()),
         lexical_class = lexical_class %>%
           fct_recode("Verbs" = "verbs", "Adjectives" = "adjectives", "Function Words" = "function_words") %>%
           fct_relevel("Verbs", "Adjectives", "Function Words"))

ggplot(plot_pred_areas, aes(y = order, x = mean)) +
  ggstance::geom_pointrangeh(aes(xmin = ci_lower, xmax = ci_upper),
                             colour = lexcat_colours["Predicates"]) +
  facet_wrap(~lexical_class, scales = "free_y", drop = TRUE) +
  geom_vline(xintercept = 0, linetype = .refline, colour = .grey) +
  scale_y_discrete(breaks = plot_pred_areas$order,
                   labels = plot_pred_areas$langform) +
  labs(x = glue("Relative representation (understand)"), y = "") +
  theme(panel.grid.major.y = .coef_line,
        axis.text.y = element_text(size = rel(0.8)))

write.csv(plot_pred_areas, file="bias_WG_Ref_NounVocab_2025Jun18.csv", row.names=FALSE)
```



# Production (WS)

```{r}
# Oxford CDI & produces -> change its form and form_type to "WS"
df_merged <- df_merged %>%
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))

vocab_comp_data <- vocab_comp_data %>% 
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))

class_data <- class_data %>%
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))

area_summary <- area_summary %>%
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))

pred_area_summary <- pred_area_summary %>% 
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))  

noun_area_summary <- noun_area_summary %>% 
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "WS", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WS", form_type))  
```

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}

base_plot(WSs, "produces") +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)
```

```{r catsyn-plot-areas-ws, fig.cap="Relative representation in vocabulary compared to chance for nouns and predicates for production data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
plot_areas_ws <- area_summary %>%
  filter(form %in% WSs, measure == "produces") %>%
  arrange(lexical_category, mean) %>%
  mutate(order = factor(row_number()))

plot_areas_ws %>%
  filter(lexical_category != "Function words") %>%
  plot_areas("production")

plot_areas_ws_cats <- plot_areas_ws %>%
  filter(lexical_category != "Function words") %>%
  select(language, form, lexical_category, mean) %>%
  pivot_wider(names_from = lexical_category, values_from = mean)

npcor_ws <- cor.test(plot_areas_ws_cats$Nouns, plot_areas_ws_cats$Predicates)
```

```{r catsyn-function-words-prod, fig.cap="Relative representation in vocabulary compared to chance for function words for production data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
plot_areas_ws %>%
  filter(lexical_category == "Function words") %>%
  plot_areas("production")
```


## Ref = Toral vocab size

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}
noun_plot(WSs, "produces", class_data) +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
#  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)
```


## Case visualization
```{r catsyn-plot-area-demo, fig.cap="For American English WS data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", fig.height=3}
demo_lang <- "English (American)"
demo_form <- "WS"
demo_data <- class_data %>% filter(language == demo_lang, form == demo_form, measure == "produces") %>% 
  subset(lexical_class %in% c("nouns", "verbs", "adjectives"))

demo_predictions <- demo_data %>%
  group_by(lexical_class) %>%
  nest() %>%
  mutate(model = map(data, ~clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = .)),
         predictions = map(model, ~tibble(vocab = pts, prop = predict(., newdata = tibble(vocab = pts))))) %>%
  select(lexical_class, predictions) %>%
  unnest(predictions) %>%
  mutate(lexical_class = factor(lexical_class, levels = c("nouns","verbs","adjectives"),
                                labels = c("Nouns","Verbs","Adjectives")))

lex_cols <- c("Nouns" = "#2C7BB6", "Verbs" = "#1A9850", "Adjectives" = "#D95F02")

ggplot(demo_predictions, aes(x = vocab, y = prop, colour = lexical_class)) +
  coord_fixed() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.75, color = "grey35") +
  geom_line(linewidth = 1.2) +
  scale_colour_manual(values = lex_cols, name = NULL) +
  vocab_scale +
  prop_scale +
  labs(title = demo_lang) +
  theme_bw() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.text = element_text(size = 10),
    legend.key.width = unit(12, "pt"),
    strip.background = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank())
```


```{r catsyn-plot-area-demo, fig.cap="For American English WS data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", fig.height=3}
demo_lang <- "English (American)"
demo_form <- "WS"
pts <- seq(0, 1, 0.01)

demo_data <- class_data %>% 
  filter(language == demo_lang, form == demo_form, measure == "produces") %>% 
  filter(lexical_class %in% c("nouns", "verbs", "adjectives")) %>%
  mutate(lexical_class = factor(lexical_class, levels = c("nouns","verbs","adjectives"),
                                labels = c("Nouns","Verbs","Adjectives")))

demo_predictions <- demo_data %>%
  group_by(lexical_class) %>%
  nest() %>%
  mutate(model = map(data, ~clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = .)),
         predictions = map(model, ~tibble(vocab = pts, prop = predict(., newdata = tibble(vocab = pts))))) %>%
  select(lexical_class, predictions) %>%
  unnest(predictions)

demo_poly <- demo_predictions %>%
  group_by(lexical_class) %>%
  group_modify(~{
    dd <- .x %>% arrange(vocab)
    diag <- tibble(vocab = rev(dd$vocab), prop = rev(dd$vocab))
    bind_rows(dd %>% select(vocab, prop), diag)
  }) %>%
  ungroup()

lex_cols <- c("Nouns" = "#2C7BB6", "Verbs" = "#1A9850", "Adjectives" = "#D95F02")

vocab_scale <- scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                  breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                  name = "Total vocabulary size (proportion)")
prop_scale <- scale_y_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                 breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                 name = "Proportion of category")

ggplot() +
  facet_wrap(~ lexical_class, nrow = 1) +
  coord_fixed() +
  geom_point(data = demo_data, aes(x = vocab, y = prop, colour = lexical_class), size = 0.25, alpha = 0.02) +
  geom_polygon(data = demo_poly, aes(x = vocab, y = prop, fill = lexical_class),
               alpha = 0.4, colour = NA) +
  geom_line(data = demo_predictions, aes(x = vocab, y = prop, colour = lexical_class),
            linewidth = 1.2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.75, color = "grey35") +
  scale_colour_manual(values = lex_cols, guide = "none") +
  scale_fill_manual(values = lex_cols, guide = "none") +
  vocab_scale +
  prop_scale +
  labs(title = demo_lang) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_rect(fill = "grey90", color = NA),
    plot.title = element_text(size = 11, hjust = 0.5),
    panel.grid.major.y = element_line(linewidth = 0.15, color = "grey85"),
    panel.grid.major.x = element_line(linewidth = 0.25, color = "grey87"),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 8, color="gray15"),
    axis.title.y = element_text(size = 10, color="gray15", margin = margin(r = 2)),
    axis.title.x = element_text(size = 10, color="gray15", margin = margin(t = 2)),
    axis.ticks = element_blank(),
    legend.position = "none",
    plot.margin = margin(2, 2, 2, 2))
ggsave("Fig0_demo.png", width = 4.8, height = 2, dpi = 320, bg = "white")
```


```{r catsyn-plot-area-demo, fig.cap="For American English WS data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", fig.height=3}
demo_langs <- c("English (American)", "Mandarin (Beijing)", "Japanese")
demo_form <- "WS"
pts <- seq(0, 1, 0.01)

demo_data3 <- class_data %>%
  filter(language %in% demo_langs, form == demo_form, measure == "produces") %>%
  filter(lexical_class %in% c("nouns", "verbs", "adjectives")) %>%
  mutate(
    lexical_class = factor(lexical_class, levels = c("nouns","verbs","adjectives"),
                           labels = c("Nouns","Verbs","Adjectives")),
    language = factor(language, levels = demo_langs))

demo_predictions3 <- demo_data3 %>%
  group_by(language, lexical_class) %>%
  nest() %>%
  mutate(model = map(data, ~clm(prop ~ I(vocab ^ 3) + I(vocab ^ 2) + vocab - 1, data = .)),
         predictions = map(model, ~tibble(vocab = pts, prop = predict(., newdata = tibble(vocab = pts))))) %>%
  select(language, lexical_class, predictions) %>%
  unnest(predictions)

demo_area3 <- demo_predictions3 %>%
  mutate(diff = prop - vocab) %>%
  group_by(language, lexical_class) %>%
  summarise(area = sum(abs(diff)) * (pts[2] - pts[1]), .groups = "drop")

demo_poly3 <- demo_predictions3 %>%
  group_by(language, lexical_class) %>%
  group_modify(~{
    dd <- .x %>% arrange(vocab)
    diag <- tibble(vocab = rev(dd$vocab), prop = rev(dd$vocab))
    bind_rows(dd %>% select(vocab, prop), diag)
  }) %>%
  ungroup()

lex_cols <- c("Nouns" = "#2C7BB6", "Verbs" = "#1A9850", "Adjectives" = "#D95F02")

vocab_scale <- scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                  breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                  name = "Vocabulary proportion")
prop_scale <- scale_y_continuous(limits = c(0, 1), expand = c(0.01, 0),
                                 breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1),
                                 name = "Proportion of category")

ggplot(demo_predictions3, aes(x = vocab, y = prop, colour = lexical_class)) +
  facet_wrap(~ language, nrow = 1) +
  coord_fixed() +
  geom_polygon(data = demo_poly3, aes(x = vocab, y = prop, fill = lexical_class),
               alpha = 0.12, colour = NA) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 0.75, color = "grey35") +
  geom_line(linewidth = 1.2) +
  scale_colour_manual(values = lex_cols, name = NULL) +
  scale_fill_manual(values = lex_cols, guide = "none") +
  vocab_scale +
  prop_scale +
  theme_bw() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.text = element_text(size = 10, color = "gray15"),
    legend.key.width = unit(12, "pt"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "grey90", color = NA),
    panel.grid.major.y = element_line(linewidth = 0.15, color = "grey85"),
    panel.grid.major.x = element_line(linewidth = 0.25, color = "grey87"),
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 10, color="gray15"),
    axis.title.y = element_text(size = 12, color="gray15", margin = margin(r = 2)),
    axis.title.x = element_text(size = 12, color="gray15", margin = margin(t = 2)),
    axis.ticks = element_blank(),
    plot.margin = margin(4, 4, 4, 4)
  )
```


## Ref = Noun vocab size

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}
noun_plot(WSs, "produces", df_merged) +
  geom_abline(slope = 1, intercept = 0, colour = .grey, linetype = .refline) +
#  geom_jitter(size = .6, alpha = .1) +
  geom_smooth(method = "clm", formula = y ~ I(x ^ 3) + I(x ^ 2) + x - 1,
              size = 1, se = FALSE)
```

## Bias strength (Ref = Total vocab size)

```{r catsyn-plot-areas-preds, fig.cap="Relative representation in vocabulary compared to chance for verbs and adjectives for production data in each language (line ranges indicate bootstrapped 95\\% confidence intervals)."}
plot_pred_areas <- pred_area_summary %>%
  filter(form_type == "WS", measure == "produces",
         lexical_class %in% c("verbs", "adjectives", "function_words", "nouns")) %>%
  ungroup() %>%
  arrange(lexical_class, mean) %>%
  mutate(order = factor(row_number()),
         lexical_class = lexical_class %>%
           fct_recode("Verbs" = "verbs", "Adjectives" = "adjectives", "Function Words" = "function_words", "Nouns" = "nouns") %>%
           fct_relevel("Nouns", "Verbs", "Adjectives", "Function Words"))

ggplot(plot_pred_areas, aes(y = order, x = mean)) +
  ggstance::geom_pointrangeh(aes(xmin = ci_lower, xmax = ci_upper),
                             colour = lexcat_colours["Predicates"]) +
  facet_wrap(~lexical_class, scales = "free_y", drop = TRUE, nrow=1) +
  geom_vline(xintercept = 0, linetype = .refline, colour = .grey) +
  scale_y_discrete(breaks = plot_pred_areas$order,
                   labels = plot_pred_areas$langform) +
  labs(x = glue("Relative representation (production)"), y = "") +
  theme(panel.grid.major.y = .coef_line,
        axis.text.y = element_text(size = rel(0.8)))


# Oxford CDI: Get back
plot_pred_areas %>% filter(language=="English (British)") %>% select(language, form, form_type, langform) %>% distinct()

plot_pred_areas <- plot_pred_areas %>% 
  mutate(form = if_else(language == "English (British)" & langform == "English (British) OxfordShort", "WSShort", form))


write.csv(plot_pred_areas, file="bias_WS_Ref_TotalVocab_2025Jun18.csv", row.names=FALSE)
```


## Bias strength (Ref = Total vocab size)

```{r catsyn-plot-points-ws, fig.asp=1.1, fig.cap="For each language's production data, proportion of each lexical category produced by each child as a function of the proportion of all vocabulary items produced by that child. Lines show model fits.", dependson="catsyn-base-plot"}

plot_pred_areas <- noun_area_summary %>%
  filter(form_type == "WS", measure == "produces",
         lexical_class %in% c("verbs", "adjectives", "function_words")) %>%
  ungroup() %>%
  arrange(lexical_class, mean) %>%
  mutate(order = factor(row_number()),
         lexical_class = lexical_class %>%
           fct_recode("Verbs" = "verbs", "Adjectives" = "adjectives", "Function Words" = "function_words") %>%
           fct_relevel("Verbs", "Adjectives", "Function Words"))

ggplot(plot_pred_areas, aes(y = order, x = mean)) +
  ggstance::geom_pointrangeh(aes(xmin = ci_lower, xmax = ci_upper),
                             colour = lexcat_colours["Predicates"]) +
  facet_wrap(~lexical_class, scales = "free_y", drop = TRUE) +
  geom_vline(xintercept = 0, linetype = .refline, colour = .grey) +
  scale_y_discrete(breaks = plot_pred_areas$order,
                   labels = plot_pred_areas$langform) +
  labs(x = glue("Relative representation (production)"), y = "") +
  theme(panel.grid.major.y = .coef_line,
        axis.text.y = element_text(size = rel(0.8)))


# Oxford CDI: Get back
plot_pred_areas %>% filter(language=="English (British)") %>% select(language, form, form_type, langform) %>% distinct()

plot_pred_areas <- plot_pred_areas %>% 
  mutate(form = if_else(language == "English (British)" & langform == "English (British) OxfordShort", "WSShort", form))


write.csv(plot_pred_areas, file="bias_WS_Ref_NounVocab_2025Jun18.csv", row.names=FALSE)
```


# Admin summary

```{r}
admin_master <- class_data %>% select(language, form, form_type, data_id, measure) %>% distinct() %>%
  mutate(
    form = if_else(language == "English (British)" & measure == "produces", "Oxford CDI", form),
    form_type = if_else(language == "English (British)" & measure == "produces", "WG", form_type)) %>% 
  left_join(admin_data)

admin_summary <- admin_master %>% group_by(language, form, form_type, measure) %>% 
  summarize(N.admin = n(), 
            N.male = sum(sex=="Male", na.rm=TRUE),
            N.female = sum(sex=="Female", na.rm=TRUE),
            N.sexunknown = N.admin - N.male - N.female,
            M.age = mean(age, na.rm=TRUE), SD.age = sd(age, na.rm=TRUE), 
            Min.age = min (age, na.rm=TRUE), Max.age = max(age, na.rm=TRUE))

write.csv(admin_summary, file="admin_summary_2025Jun18.csv", row.names=FALSE)
```

